Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2020-09-30 manuscript/ERL_global_C_review.Rmd
cleanup

**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

@@

Becky Banbury Morgan^3^

@@

#ForC simplified
ForC_simplified <- read.csv(paste0(path_to_ForC, "ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)

n_no_history <- sum(ForC_simplified[,"history.no.info"]) #(relative to ForC_simplified)
n_managed_in_ForC_simplified <- sum(ForC_simplified[,"managed"])
n_disturbed_in_ForC_simplified <- sum(ForC_simplified[,"disturbed"])
records_eligible_for_inclusion <- sum(ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history

@@

*NOTES TO COAUTHORS: *

- "**???**" indicates a reference that we have entered in the .Rmd file, but not yet in the bibliography file. Don't worry about those. (However, places with "REF" need references)
- bold/ italic text indicates places where I know we need to go back and revisit/ fill in info

@@

*Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we use `r n_records` *ForC* records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables.

@@

*Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation. The synthetic and internally consistent overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.

@@

Forest ecosystems are shaping the course of climate change [@ipcc_global_2018] through their influence on atmospheric carbon dioxide (CO~2~).
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.

The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change. 
Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important.
Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
Yet, this sink is reduced by deforestation, estimated at 3.7 Gt CO~2~ yr-1 ranging from 1990-2020
Recent net deforestation (*i.e.*, gross deforestation minus regrowth) has been a source of CO~2~ emissions, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].

Despite the importance of forests, comprehensive global studies have historically been limited by the scattered and more local nature of research studies. 
Primary research articles typically cover only a small numbers of sites at a time.
The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015
], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*.
In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
*ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration). 
These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
These data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
Since the its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].

@@

This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016, @anderson-teixeira_forc_2018].
*ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v.2.0* [@anderson-teixeira_forc_2018].
Since publication of ForC v.2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v.##, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
We've also added data from individual publications (detailed list at https://github.com/forc-db/ForC/blob/master/database_management_records/ForC_data_additions_log.csv), with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).

To facilitate analyses, we created a simplified version of ForC, *ForC-simplified* (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which we analyzed here.
In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
Duplicate or otherwise conflicting records were reconciled as described in *APPENDIX S1*, resulting in a total of `r nrow(ForC_simplified)` records (`r round(nrow(ForC_simplified)/nrow(ForC_measurements)*100,digits=1)`% size of total database).
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event.
Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%).
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/nrow(ForC_simplified)*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/nrow(ForC_simplified)*100,digits=1)`%).
In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.

@@

To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R package version **x.xx**; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
This analysis was run for variables with records for at least seven distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).

@@

To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R (**version, citation**).

@@

Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met our strict criteria for inclusion in this study (Fig. 1).
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ **3** distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.

@@

Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs. 7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
Unfortunately, data remain too sparse for statistical comparison across biomes (Figs. 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2005]. 
Further research on non-living C stocks in the world's forests will be essential to completing the picture.

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig. 6), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2020; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.

@@

In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_global_2011]--and more variable age trends in deadwood and $OL$.
The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig. 7e; e.g., @vargas_biomass_2008].
In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig. 7e; e.g., @carmona_coarse_2002].
Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.

>>>>

**Title:** Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the Global Forest Carbon (ForC) database

@@

Becky Banbury Morgan^1,3^

@@

#ForC simplified
ForC_simplified <- read.csv(paste0(path_to_ForC, "ForC_simplified/ForC_simplified.csv"), stringsAsFactors = FALSE, check.names = FALSE)
ForC_simplified_duplicates_removed <- sum(ForC_simplified$suspected.duplicate ==0)

n_no_history <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$history.no.info ==1) #(relative to ForC_simplified)
n_managed_in_ForC_simplified <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$managed ==1)
n_disturbed_in_ForC_simplified <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$disturbed ==1)
records_eligible_for_inclusion <- sum(ForC_simplified$suspected.duplicate ==0 & ForC_simplified$managed ==0 & ForC_simplified$disturbed ==0 & ForC_simplified$history.no.info == 0)  #number in ForC_simplified that are not disturbed, managed, or no.history

@@

@@

*Methods/Design.* Here, we draw from the Global Forest Carbon Database, ForC, to provide a macroscopic overview of C cycling in the world’s forests, giving special attention to stand age-related variation. Specifically, we use `r n_records` *ForC* records from `r n_areas_tot` geographic locations representing `r nrow(T1)` C cycle variables to characterize ensemble C budgets for four broad forest types (tropical broadleaf evergreen, temperate broadleaf, temperate conifer, and taiga), including estimates for both mature and regrowth (age <100 years) forests. For regrowth forests, we quantify age trends for all variables with sufficient data.

@@

*Discussion.* As climate change accelerates, understanding and managing the carbon dynamics of forests is critical to forecasting, mitigation, and adaptation. This synthetic and internally consistent global overview of C stocks and fluxes across biomes and stand ages will help to advance these efforts.

@@

Forest ecosystems are shaping the course of climate change through their influence on atmospheric carbon dioxide [CO~2~; @bonan_forests_2008; @friedlingstein_global_2019; @ipcc_global_2018].
Despite the centrality of forest C cycling in regulating atmospheric CO~2~, important uncertainties in climate models [@friedlingstein_climatecarbon_2006; @krause_large_2018; @bonan_model_2019; @di_vittorio_initial_2020] and CO~2~ accounting frameworks [@pan_large_2011] can be traced to lack of accessible, comprehensive data on how C cycling varies across forest types and in relation to stand history.
These require large-scale databases with global coverage, which runs contrary to the nature in which forest C stocks and fluxes are measured and published.
While remote sensing measurements are increasingly useful for global- or regional-scale estimates of a few critical variables [e.g., aboveground biomass: @hu_mapping_2016; @spawn_harmonized_2020; gross primary productivity, $GPP$: @li_mapping_2019; @saatchi_benchmark_2011], measurement and validation of most forest C stocks and fluxes require intensive on-the-ground data collection.

A robust understanding of forest impacts on global C cycling is essential. 
Annual gross CO~2~ sequestration in forests (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ [@badgley_terrestrial_2019], or >7 times average annual fossil fuel emissions from 2009-2018 [9.5 $\pm$ 0.5 Gt C yr^-1^; @friedlingstein_global_2019].
Most of this enormous C sequestration is counterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, with forests globally dominant as sources of both soil respiration [@warner_spatial_2019] and fire emissions [@van_der_werf_global_2017].
In recent years, the remaining CO~2~ sink averaged 3.2 $\pm$ 0.6 GtC yr^-1^ from 2009-2018, offsetting 29% of anthropogenic fossil fuel emissions [@friedlingstein_global_2019].
Yet, this sink is reduced by deforestation, estimated at ~1 Gt C yr^-1^ in recent decades [@pan_large_2011; @tubiello_carbon_2020], reducing the net forest sink to ~1.1-2.2 Gt C yr^-1^ across Earth's forests [@friedlingstein_global_2019].

The future of the current forest C sink is dependent both upon forest responses to climate change itself and human land use decisions, which will feedback and strongly influence the course of climate change. 
Regrowing forests in particular will play an important role [@pugh_role_2019], as almost two-thirds of the world’s forests were secondary as of 2010 [@fao_global_2010].
As anthropogenic and climate-driven disturbances impact an growing proportion of Earth's forests [@andela_human-driven_2017; @mcdowell_pervasive_2020], understanding the carbon dynamics of regrowth forests is increasingly important [@andersonteixeira_altered_2013].
Although age trends in aboveground biomass have been relatively well-studied and synthesized globally [@cook-patton_mapping_2020], a relative dearth of data and synthesis on other C stocks and fluxes in secondary forests points to an under-filled need to characterize age-related trends in forest C cycling.
Such understanding is particularly critical for reducing uncertainty regarding the potential for carbon uptake and climate change mitigation by regrowth forests [@krause_large_2018; @cook-patton_mapping_2020].
Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thus central to efforts to mitigate climate change [@grassi_key_2017; @griscom_natural_2017; @cavaleri_urgent_2015].

Despite the importance of forests, comprehensive global studies have historically been limited by the scattered and more local nature of research studies. 
Primary research articles typically cover only a small numbers of sites at a time.
The rare exceptions that span regions or continents with rare exceptions spanning regions or continents are typically coordinated through research networks such as ForestGEO [@anderson-teixeira_ctfs-forestgeo_2015, e.g., @lutz_global_2018], NEON [@schimel_neon_2007], or FLUXNET [@baldocchi_fluxnet_2001; e.g., @novick_ameriflux_2018].
The result of decades of research on forest C cycling is that tens of thousands of records have been distributed across literally thousands of scientific articles --often behind paywalls-- along with variation in data formats, units, measurement methods, *etc.*
In this format, the data are effectively inaccessible for many global-scale analyses, including those attempting to benchmark model performance with global data (Clark et al 2017, Luo et al 2012), quantify the the role of forests in the global C cycle [*e.g.*, @pan_large_2011], or use book-keeping methods to quantify actual or scenario-based exchanges of CO~2~ between forests and the atmosphere [@griscom_natural_2017; @houghton_terrestrial_2020].

To address the need for global-scale analyses of forest C cycling, we recently developed *ForC* [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
*ForC* contains published estimates of forest ecosystem C stocks and annual fluxes (>50 variables), with the different variables capturing the unique ecosystem pools (e.g., woody, foliage, and root biomass; dead wood) and flux types (e.g, gross and net primary productivity; soil, root, and ecosystem respiration). 
These data are ground-based measurements, and *ForC* contains associated data required for interpretation (*e.g.*, stand history, measurement methods).
Data have been amalgamated from original peer-reviewed publications, either directly or via intermediary data compilations.
Since the its most recent publication [@anderson-teixeira_forc_2018], *ForC* has grown to include two additional large databases: the Global Soil Respiration Database [SRDB; @bond-lamberty_global_2010] and the Global Reforestation Opportunity Assessment database [GROA; @cook-patton_mapping_2020] that also synthesized published forest C data.
Following these additions, *ForC* currently contains `r nrow(ForC_measurements)` records from `r nrow(ForC_plots)` plots and `r n_areas` distinct geographic areas representing all forested biogeographic and climate zones.
This represents an `r round((nrow(ForC_measurements) - 17367)/17367*100, digits=0)`% increase in records from the prior publication [@anderson-teixeira_forc_2018].

@@

This review synthesizes data from the *ForC* database [Fig. 1; https://github.com/forc-db/ForC; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
*ForC* amalgamates numerous intermediary data sets [*e.g.*, @luyssaert_co2_2007; @bond-lamberty_global_2010; @cook-patton_mapping_2020] and original studies.
Original publications were referenced to check values and obtain information not contained in intermediary data sets, although this process has not been completed for all records.
The database was developed with goals of understanding how C cycling in forests varies across broad geographic scales and as a function of stand age.
As such, there has been a focus on incorporating data from regrowth forests (*e.g.*, Anderson et al 2006, Martin et al 2013, Bonner et al 2013) and obtaining stand age data when possible [83% of records in v.2.0; @anderson-teixeira_forc_2018].
Particular attention was given to developing the database for tropical forests [@anderson-teixeira_carbon_2016], which represented roughly one-third of records in *ForC* v2.0 [@anderson-teixeira_forc_2018].
Since publication of ForC v2.0, we added the following data to ForC: the Global Database of Soil Respiration Database [*SRDB* v4, `r n_SRDB` records; @bond-lamberty_global_2010], the Global Reforestation Opportunity Assessment database [*GROA* v1.0, `r n_GROA` records; @cook-patton_mapping_2020; @anderson-teixeira_forc-dbgroa_2020].
We have also added data from individual publications, with a particular focus on productivity [@banbury_morgan_global_nodate; e.g., @taylor_temperature_2017], dead wood, and ForestGEO sites [e.g., @lutz_global_2018, @johnson_climate_2018].
The database version used for this analysis has been tagged as a new release on Github (v3.0) and assigned a DOI through Zenodo (DOI: TBD).

To facilitate analyses, we created a simplified version of ForC, *ForC-simplified* (https://github.com/forc-db/ForC/blob/master/ForC_simplified), which we analyzed here.
In generating *ForC-simplified*, all measurements originally expressed in units of dry organic matter ($OM$) were converted to units of C using the IPCC default of $C = 0.47 * OM$ [@ipcc_global_2018].
Duplicate or otherwise conflicting records were reconciled as described in Appendix S1, resulting in a total of `r ForC_simplified_duplicates_removed` records (`r round(ForC_simplified_duplicates_removed/nrow(ForC_measurements)*100,digits=1)`% size of total database).
Records were filtered to remove plots that had undergone significant anthropogenic management or major disturbance since the most recent stand initiation event.
Specifically, we removed all plots flagged as managed in ForC-simplified (`r round(n_managed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%).
This included plots with any record of managements manipulating CO~2~, temperature, hydrology, nutrients, or biota, as well as any plots whose site or plot name contained the terms “plantation”, "planted", “managed”, “irrigated”, or “fertilized”.
Plots flagged as disturbed in ForC-simplified (`r round(n_disturbed_in_ForC_simplified/ForC_simplified_duplicates_removed*100,digits=1)`%) included stands that had undergone any notable anthropogenic thinning or partial harvest.
We retained sites that were grazed or had undergone low severity natural disturbances (<10% mortality) including droughts, major storms, fires, and floods.
We removed all plots for which no stand history information had been retrieved (`r round(n_no_history/ForC_simplified_duplicates_removed*100,digits=1)`%).
In total, this resulted in `r records_eligible_for_inclusion` records (`r round(records_eligible_for_inclusion/nrow(ForC_measurements)*100,digits=1)`% of the records in the database) being eligible for inclusion in the analysis.

@@

To test for differences across mature forest biomes, we also examined how stand age impacted fluxes and stocks, employing a mixed effects model ['lmer' function in 'lme4' R; @bates_fitting_2015] with biome as fixed effect and plot nested within geographic.area as random effects on the intercept.
When Biome had a significant effect, we looked at a Tukey's pairwise comparison to see which biomes were significantly different from one another. 
This analysis was run for variables with records for at least seven distinct geographic areas in more than one biome, excluding any biomes that failed this criteria (Table 1).

@@

To facilitate the accessibility of our results and data, and to allow for rapid updates as additional data become available, we have automated all database manipulation, analyses, and figure production in R [@r_core_team_r_2020].

@@

Of the `r nrow(ForC_measurements)` records in *ForC* v3.0, `r n_records` met our strict criteria for inclusion in this study (Fig. 1).
These records were distributed across `r n_plots_tot` plots in `r n_areas_tot` distinct geographic areas.
Of the 23 flux and 11 stock variables mapped in these diagrams, *ForC* contained sufficient mature forest data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 7 distinct geographic areas) for `r n_fluxes_with_ge7_areas[["Tropical broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Tropical broadleaf MATURE"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate broadleaf MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate broadleaf MATURE"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge7_areas[["Temperate conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Temperate conifer MATURE"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge7_areas[["Boreal conifer MATURE"]]` fluxes and `r n_stocks_with_ge7_areas[["Boreal conifer MATURE"]]` stocks in boreal forests.
For regrowth forests (<100 yrs), *ForC* contained sufficient data for inclusion in our statistical analyses (*i.e.*, records from $\ge$ 3 distinct geographic areas) for `r n_fluxes_with_ge3_areas[["Tropical broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Tropical broadleaf YOUNG"]]` stocks in tropical broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate broadleaf YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate broadleaf YOUNG"]]` stocks in temperate broadleaf forests, `r n_fluxes_with_ge3_areas[["Temperate conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Temperate conifer YOUNG"]]` stocks in temperate conifer forests, and `r n_fluxes_with_ge3_areas[["Boreal conifer YOUNG"]]` fluxes and `r n_stocks_with_ge3_areas[["Boreal conifer YOUNG"]]` stocks in boreal forests.

@@

Whereas biomass can be remotely sensed and receives significant research attention, far less is known about geographical variation in deadwood and organic layer ($OL$) across biomes, which has proved a limitation for C accounting efforts [@pan_large_2011].
Although these stocks can be important--exceeding 100 t C ha^-1^ in some stands (Figs. 7d-e, S23-S25), this study is the first to synthesize deadwood data on a global scale [but see @cook-patton_mapping_2020 for young forests].
Unfortunately, data remain too sparse for statistical comparison across biomes (Figs. 7, S23-S25; but see below for age trends), pointing to a need for more widespread quantification of both standing and downed deadwood.
*ForC* coverage of $OL$ stocks is more comprehensive, revealing no significant differences across temperate and tropical biomes, but a tendency towards higher $OL$ in boreal forests, consistent with the idea that slower decomposition in colder climates results in more buildup of organic matter [@allen_global_2002]. 
Further research on non-living C stocks in the world's forests will be essential to completing the picture.

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Fig. 6), consistent with current understanding of age trends in forest C cycling [e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_auto$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. 7).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.

@@

In terms of stocks, our study reveals consistent increases in live biomass stocks with stand age--a pattern that is well-known and expected [e.g., @lichstein_biomass_2009; @yang_carbon_2011]--and more variable age trends in deadwood and $OL$.
The latter are particularly sensitive to the type of disturbance, where disturbances that remove most organic material (e.g., logging, agriculture) result in negligible deadwood in young stands, followed by a buildup over time [tropical stands in Fig. 7e; e.g., @vargas_biomass_2008].
In contrast, natural disturbances (e.g., fire, drought) can produce large amounts of deadwood (mostly $DW_{standing}$) that slowly decomposes as the stand recovers, resulting in declines across young stand ages [e.g., temperate and boreal stands in Fig. 7e; e.g., @carmona_coarse_2002].
Again, further study and synthesis of non-living C stocks across biomes and stand ages will be valuable to giving a more comprehensive picture.
